Data Processing Strategies to Determine Maximum Oxygen Uptake: A Systematic Scoping Review and Experimental Comparison with Guidelines for Reporting

### 2023-07-27 scripts/comparison.Rmd
second revision work

# Line plots with large data are very slow in ggplot2
# using a scatter plot instead
p_rr <- ggplot(rr_all, aes(x = time, y = value_rr, colour = is_end, alpha = is_end)) +
  geom_point(size = 0.5, show.legend = FALSE) +
  scale_colour_manual(values = c("black", met.brewer("Peru2")[2])) +
  scale_alpha_manual(values = c(0.05, 0.3)) +
  scale_x_continuous(
    name = NULL,
    limits = c(60,950),
    breaks = c(180, 360, 540, 720, 900),
    labels = c(2, 5, 8, 12, 15),
    expand = c(0,0)
  ) +
  scale_y_continuous(
    name = expression(paste("Respiratory Rate ", (min^-1))),
    breaks = c(20,40,60,80)
  ) +
  theme_classic()

p_vo2 <- ggplot(rr_all, aes(x = time, y = value_vo2, colour = is_end, alpha = is_end, group = id)) +
  geom_point(size = 0.5, show.legend = FALSE) +
  scale_colour_manual(values = c("black", met.brewer("Peru2")[2])) +
  scale_alpha_manual(values = c(0.05, 0.3)) +
  scale_x_continuous(
    name = "Time (min)",
    limits = c(60,950),
    breaks = c(180, 360, 540, 720, 900),
    labels = c(2, 5, 8, 12, 15),
    expand = c(0,0)
  ) +
  scale_y_continuous(
    name = expression(paste(dot(V), O[2], " uptake ", (ml%.%min^-1%.%kg^-1)))
  ) +
  theme_classic()

p_all <- p_rr / p_vo2 + plot_annotation(tag_levels = 'A')

@@

p_all_vert <- (p_rr | p_vo2) + plot_annotation(tag_levels = 'A')
# ggsave(filename = "../plots/timecourse_vert.png", plot = p_all_vert, width = 8, height = 3, dpi = 600, bg = "white")

@@

p_selcomp <- ggplot(slct_long, aes(x = value, y = name)) +
  geom_line(aes(group = id), orientation = "y", colour = "grey80") +
  stat_dotsinterval(
    alpha = 0.7, 
    .width = c(0.5, 0.8), 
    binwidth = 0.5, dotsize = 1,
    point_interval = "mean_qi"
    ) +
  scale_x_continuous(
    name = expression(paste(dot(V), O[2][max], " ", (ml%.%min^-1%.%kg^-1))),
    expand = c(0.02,0.02),
    breaks = seq(50,75,5),
    labels = c("50","","60","","70","")
  ) +
  scale_y_discrete(
    name = NULL,
    expand = c(0.05,0.1)
  ) +
  theme_classic(base_size = 14) +
  theme(
    axis.line.y = element_blank(),
    axis.text.y = element_text(colour = "black")
  )

@@

p_selcomp
```

>>>>

# Line plots with large data are very slow in ggplot2
# using a scatter plot instead
p_rr <- ggplot(rr_all, aes(x = time, y = value_rr, colour = is_end, alpha = is_end)) +
  geom_point(size = 0.5, show.legend = FALSE) +
  scale_colour_manual(values = c("black", met.brewer("Peru2")[2])) +
  scale_alpha_manual(values = c(0.05, 0.3)) +
  scale_x_continuous(
    name = NULL,
    limits = c(60,950),
    breaks = c(180, 360, 540, 720, 900),
    labels = c(2, 5, 8, 11, 14),
    expand = c(0,0)
  ) +
  scale_y_continuous(
    name = expression(paste("Respiratory Rate ", (min^-1))),
    breaks = c(20,40,60,80)
  ) +
  theme_classic()

p_vo2 <- ggplot(rr_all, aes(x = time, y = value_vo2, colour = is_end, alpha = is_end, group = id)) +
  geom_point(size = 0.5, show.legend = FALSE) +
  scale_colour_manual(values = c("black", met.brewer("Peru2")[2])) +
  scale_alpha_manual(values = c(0.05, 0.3)) +
  scale_x_continuous(
    name = "Time (min)",
    limits = c(60,950),
    breaks = c(180, 360, 540, 720, 900),
    labels = c(2, 5, 8, 11, 14),
    expand = c(0,0)
  ) +
  scale_y_continuous(
    name = expression(paste(dot(V), O[2], " uptake ", (ml%.%min^-1%.%kg^-1)))
  ) +
  theme_classic()

p_all <- p_rr / p_vo2 + plot_annotation(tag_levels = 'a')

@@

p_all_vert <- (p_rr | p_vo2) + plot_annotation(tag_levels = 'a')
# ggsave(filename = "../plots/timecourse_vert.png", plot = p_all_vert, width = 8, height = 3, dpi = 600, bg = "white")

@@

# plot with individual data
p_selcomp <- ggplot(slct_long, aes(x = value, y = name)) +
  geom_line(aes(group = id), orientation = "y", colour = "grey80") +
  stat_dotsinterval(
    alpha = 0.7, 
    .width = c(0.33, 0.66), 
    binwidth = 0.5, dotsize = 1,
    point_interval = "mean_qi"
    ) +
  scale_x_continuous(
    name = expression(paste(dot(V), O[2][max], " ", (ml%.%min^-1%.%kg^-1))),
    expand = c(0.02,0.02),
    breaks = seq(50,75,5),
    labels = c("50","","60","","70","")
  ) +
  scale_y_discrete(
    name = NULL,
    expand = c(0.05,0.1)
  ) +
  theme_classic(base_size = 14) +
  theme(
    axis.line.y = element_blank(),
    axis.text.y = element_text(colour = "black")
  )

@@

p_selcomp

# calculate place for significance symbol
get_mean_plus <- function(i) {
  mean(slct_long$value[slct_long$name == slct_long$name[i]], 0.9) + 6.5
}
slct_long$place <- map_dbl(seq_len(nrow(slct_long)), get_mean_plus)

get_symbol <- function(i) {
  ifelse(
    slct_long$name[i] %in% c("(d) 30-s moving", "(c) digital filter"),"#","*"
  )
}
get_vjust <- function(i) {
  ifelse(
    slct_long$name[i] %in% c("(d) 30-s moving", "(c) digital filter"), 0.3 ,0.6
  )
}
slct_long$symbol <- map_chr(seq_len(nrow(slct_long)), get_symbol)
slct_long$vjust <- map_dbl(seq_len(nrow(slct_long)), get_vjust)
# plot without individual data
p_selcomp2 <- ggplot(slct_long, aes(x = value, y = name)) +
  stat_pointinterval(
    alpha = 0.7, 
    .width = c(0.33, 0.66), 
    point_interval = "mean_qi"
  ) +
  geom_text(
    aes(x = place, label = symbol, vjust = vjust), 
    show.legend = FALSE
  ) +
  scale_x_continuous(
    name = expression(paste(dot(V), O[2][max], " ", (ml%.%min^-1%.%kg^-1))),
    expand = c(0.02,0.02),
    breaks = seq(52.5,72.5,2.5),
    labels = c("","55","","60","","65","","70","")
  ) +
  scale_y_discrete(
    name = NULL,
    expand = c(0.05,0.1)
  ) +
  theme_classic(base_size = 14) +
  theme(
    axis.line.y = element_blank(),
    axis.text.y = element_text(colour = "black")
  )

p_selcomp2

# ggsave(filename = "../plots/sel_comp2.pdf", width = 7, height = 3, dpi = 600, bg = "white")
# ggsave(filename = "../plots/sel_comp2.png", width = 7, height = 3.5, dpi = 600, bg = "white")

```

## Compare impact of binning preprocessing on VO2max

```{r}
# normal time binned
max_time_binned <- function(data, parameter, max = TRUE) {
  nbin <- ceiling(nrow(data) / parameter)
  ibins <- seq(from = 1, by = parameter, length.out = nbin)
  binmean <- function(i, length, data) {
    values <- data$VO2_rel[i:(i + length - 1)]
    values[is.na(values)] <- 0
    mean(values)
  }
  binmeans <- purrr::map_dbl(.x = ibins, .f = binmean, length = parameter, data = data)
  if (max) max(binmeans) else binmeans
}

max_time_binned_late <- function(data, parameter, max = TRUE) {
  raw <- attr(data, "raw")
  last_load <- round(raw$time[max(which(raw$load != 0))])
  nbin <- ceiling(last_load / parameter)
  ibins <- seq(to = last_load, by = parameter, length.out = nbin)
  binmean <- function(i, length, data) {
    values <- data$VO2_rel[i:(i + length - 1)]
    values[is.na(values)] <- 0
    mean(values)
  }
  binmeans <- purrr::map_dbl(.x = ibins, .f = binmean, length = parameter, data = data)
  if (max) max(binmeans) else binmeans
}

max_time_binned_noninterp <- function(data, parameter, max = TRUE) {
  raw <- attr(data, "raw")
  nbin <- ceiling(nrow(data) / parameter)
  ibins <- seq(from = 0, by = parameter, length.out = nbin)
  binmean <- function(i, length, data) {
    raw <- attr(data, "raw")
    values <- raw$VO2[raw$time >= ibins[i] & raw$time <= (ibins[i]+length)] / attr(data, "info")$weight
    values[is.na(values)] <- 0
    mean(values)
  }
  binmeans <- purrr::map_dbl(.x = seq_along(ibins), .f = binmean, length = parameter, data = data)
  if (max) max(binmeans) else binmeans
}

get_vo2max <- function(n, parameter = 30) {
  data <- spiro(files[n], weight = info$bodymass[n])
  data.frame(
    begin = max_time_binned(data = data, parameter = parameter),
    end = max_time_binned_late(data = data, parameter = parameter),
    nonint = max_time_binned_noninterp(data = data, parameter = parameter)
  )
}

o <- purrr::map_dfr(seq_along(files), get_vo2max)
o$dif_align <- o$begin - o$end
o$dif_interp <- o$begin - o$nonint
apply(o, 2, mean)
apply(o, 2, sd)

```

## Plot for article: Visualizing different strategies in one raw data set

```{r example, warning = FALSE}
plot_data <- function(id, show_legend = TRUE, show_ylabel = TRUE) {
id_info <- read.csv("../data/participants.csv")
bodymass <- id_info$bodymass[id_info$p == id]

exm_data <- spiro(files[id])
raw_vo2 <- attr(exm_data, "raw")$VO2
raw_time <- attr(exm_data, "raw")$time

fz2 <- spiro_smooth(exm_data, "fz", "VO2")[,1]
mb2 <- spiro_smooth(exm_data, "7b", "VO2")[,1]
mt2 <- spiro_smooth(exm_data, 30, "VO2")[,1]
tb2 <- rep(max_time_binned(exm_data, 60, max = FALSE) * attr(exm_data, "info")[["weight"]], each = 60)

exm_tidy2$method <- factor(
  exm_tidy2$method, 
  levels = c("raw", "mb", "mt", "tb", "fz"), 
  labels = c("raw","7-breath moving","30-s moving","60-s binned", "digital filter"),
  ordered = TRUE
)
exm_tidy2$value <- exm_tidy2$value / bodymass

allplot <- ggplot(exm_tidy_cut, aes(x = time)) +
  geom_point(
    data = exm_tidy_cut_copy, 
    aes(x = timecopy, y = value, alpha = method),
    color = "black",
    alpha = 0.15,
    show.legend = FALSE,
    size = 1.5
  ) +
  geom_line(
    data = exm_tidy_cut[exm_tidy_cut$method %in% c("7-breath moving", "30-s moving", "digital filter", "60-s binned"), ], 
    aes(y = value, color = method, linetype = method), 
    size = 0.9,
    show.legend = show_legend
  ) +
  scale_x_continuous(
    name = "Time (min)",
    limits = c(520,800),
    breaks = c(540, 660, 780),
    labels = c(8, 10, 12),
    expand = c(0,0)
  ) +
    scale_y_continuous(
      name = ifelse(show_ylabel, expression(paste(dot(V), O[2], " ", (ml%.%min^-1%.%kg^-1))), ""),
      limits = c(45,80)
    ) +
  scale_color_manual(
    name = "Method",
    values = c("darkgreen", "purple", "orange", "turquoise2")
  ) +
  scale_linetype_manual(name = "Method", values = c("solid", "dashed", "longdash", "dotted")) +
  theme_classic(12) +
  theme(legend.position = "top")

allplot
}

guide_area() / (plot_data(53, show_legend = FALSE) | plot_data(61, show_ylabel = FALSE)) + plot_layout(guides = 'collect', heights = c(1,6)) + plot_annotation(tag_levels = 'a') & theme(legend.position = "top")

#ggsave("../plots/allplot_article.png", dpi = 300, width = 7, height = 4, bg = "white")
#ggsave("../plots/allplot_article.pdf", dpi = 300, width = 7, height = 4, bg = "white")
```

### Difference between filter and 30-s moving average
```{r}
res <- read.csv("../data/results.csv")
ba <- data.frame(
  f = fmax,
  m = res$tm_30
)

ba$dif <- ba$f - ba$m

mean(ba$dif)
sd(ba$dif)
```
